---
title: ""
format: dashboard
---

```{r include=F}
#| label: setup

options(scipen=100)                         # suppress scientific notation
                                            # install pacman if needed
if (!require("pacman")) install.packages("pacman", repos = "http://lib.stat.cmu.edu/R/CRAN/")
pacman::p_load(pacman, tidyverse, ggthemes, gridExtra, magrittr,
               kableExtra, gt)

#pacman::p_loaded()

```

```{r echo=F}
#| label: created formatted date
# create variables for inline code
today <- Sys.Date()                                                # saves today's date
today_fmt <- stamp("Saturday, January 1, 1999", quiet = T)(today)  # formats date

```

# Syllabus and Course Links

## Row

**This website was last updated on `r today_fmt`.**

<br>

[**Fall 2026
Syllabus**](https://docs.google.com/document/d/1ycqRSNZWkvodKbACx_fq_Bx8cS-HBdHsLDcDNTIoz6I/edit?usp=sharing){target="_blank"}

- Syllabus is updated at the end of the most recent
  completed semester.

<br>

**Contact Information: Professor Pooler
(pspooler\@syr.edu) - Office: WSOM Room 518**

<br>

**NOTES:**

- This is the permanent version of Professor Pooler's MAS
  261 Course Website and is only updated at the end of the
  semester or when substantial changes are made to the
  material during the semester.

- Homework assignments and 'Practice Questions' for tests
  are not posted on the permanent MAS 261 Website.

- Lectures that are solely devoted to review and preparation
  for mid-semester tests and the final exam are also not
  available on the permanent website.

  - Missing lecture numbers are associated with review
    lectures and in-class test days.

[**Read this letter about
AI**](https://www.nytimes.com/2025/06/18/learning/we-need-to-chat-gpt.html?smid=url-share){target="_blank"}

- Using AI in coding and writing is commonplace and helpful
  if used cautiously and ethically.

- This is a great 'letter' from a high school student who
  'broke up' with AI because she felt it was supplanting her
  own voice. It's worth your time to read and consider it.

**Useful Software Links**

- [Posit Cloud](https://posit.cloud/){target="_blank"}

- [R](https://cran.r-project.org/){target="_blank"}

- [RStudio](https://posit.co/products/open-source/rstudio){target="_blank"}

- [Quarto](https://quarto.org/docs/get-started/){target="_blank"}

**OPTIONAL Software Training**

- [LinkedIn
  Learning](https://linkedinlearning.syr.edu/){target="_blank"} -
  Free for Syracuse University Students, Staff, and Faculty

  - [R for Data Science: Analysis and
    Visualization](https://www.linkedin.com/learning/r-for-data-science-analysis-and-visualization?u=76816226){target="_blank"}

    - This training is optional and provides a good
      introduction to R and RStudio which are used in this
      course.

  - [Python Essential
    Training](https://www.linkedin.com/learning/python-essential-training-18764650?u=76816226){target="_blank"}

    - Python is not used in this course but it is helpful
      for analytics and data science.

<br>

- [DataCamp](https://www.datacamp.com/){target="_blank"}

  - An excellent (but not free) online learning center for
    many software packages. Costs can be subsidized (see
    below).

<br>

- [Whitman WIRE
  Initiative](https://whitman.syracuse.edu/experiential-learning/wire-initiative){target="_blank"}

  - The WIRE initiative will subsidize costs of online
    software training that results in certification.

  - Contact their office to arrange a meeting and develop a
    plan.

<br>

- [SU Open Source Program Office
  (OSPO)](https://opensource.syracuse.edu/){target="_blank"}

  - OSPO offers workshops in open source software such as
    R/RStudio, Python, Github and facilitates communication
    between open source software users on campus.

# Lectures 1-16

## Row {height="45%"}

**Estimates and Confidence**

<br>

- **Lecture 16 - Proportions, Percentages and Confidence
  Intervals**

  - [**Slides**](https://rpubs.com/PeneLope_PE/1232622){target="_blank"}

  - [**Notes**](https://rpubs.com/PeneLope_PE/1232623){target="_blank"}

  - [**PDF**](https://drive.google.com/file/d/1rIuwLPW90rioY-2A-B1ksXDGg27s8wSh/view?usp=sharing){target="_blank"}

  - [**t-Table**](https://drive.google.com/file/d/1-YnYJbk-5S5AmLs1VYovg-rxnoht3hS_/view?usp=sharing){target="_blank"}

<br>

- **Lecture 14 - Effect of Sample Size and Confidence level
  on Confidence Intervals**

  - [**Slides**](https://rpubs.com/PeneLope_PE/1229841){target="_blank"}

  - [**Notes**](https://rpubs.com/PeneLope_PE/1229843){target="_blank"}

  - [**PDF**](https://drive.google.com/file/d/1QTKUEHi59h-ufQ5PVbTdssKj1wabyERf/view?usp=sharing){target="_blank"}

  - [**t-Table**](https://drive.google.com/file/d/1-YnYJbk-5S5AmLs1VYovg-rxnoht3hS_/view?usp=sharing){target="_blank"}

<br>

- **Lecture 13 - Intro. to the t Distribution and Confidence
  Intervals**

  - [**Slides**](https://rpubs.com/PeneLope_PE/1229168){target="_blank"}

  - [**Notes**](https://rpubs.com/PeneLope_PE/1229172){target="_blank"}

  - [**PDF**](https://drive.google.com/file/d/1HpmiXjRTZ4JCjx1dAbog_gWuhTHIhw6y/view?usp=sharing){target="_blank"}

<br>

**Data Distributions**

<br>

- **Lecture 9 - Central Limit Theorem**

  - [**Slides**](https://rpubs.com/PeneLope_PE/1223885){target="_blank"}

  - [**Notes**](https://rpubs.com/PeneLope_PE/1223886){target="_blank"}

  - [**PDF**](https://drive.google.com/file/d/16OV59_uT7RNe2CaQrsOsqG8s-P3A9nsm/view?usp=sharing)

<br>

- **Lecture 8 - Sampling Distribution of the Sample Mean /
  CLT**

  - [**Slides**](https://rpubs.com/PeneLope_PE/1220620){target="_blank"}

  - [**Notes**](https://rpubs.com/PeneLope_PE/1220604){target="_blank"}

  - [**PDF**](https://drive.google.com/file/d/1-8uq_NN7POxSsgKovo-i9USFuAFORYFV/view?usp=sharing){target="_blank"}

<br>

- **Lecture 7 - Empirical Rule / Finding X from a
  Probability**

  - [**Slides**](https://rpubs.com/PeneLope_PE/1220819){target="_blank"}

  - [**Notes**](https://rpubs.com/PeneLope_PE/1220821){target="_blank"}

  - [**PDF**](https://drive.google.com/file/d/1-29uIhO6YSI_jGL8472QZx99xQ1WdxH-/view?usp=sharing){target="_blank"}

<br>

- **Lecture 6 - Introduction to the Normal Distribution**

  - [**Slides**](https://rpubs.com/PeneLope_PE/1220421){target="_blank"}

  - [**Notes**](https://rpubs.com/PeneLope_PE/1217759){target="_blank"}

  - [**PDF**](https://drive.google.com/file/d/1ujIhyPile-Zu-dhTfkQD5zVdNz-T5mq2/view?usp=sharing){target="_blank"}

<br>

**Describing Data**

<br>

- **Lecture 5 - Visualizing Categorical and Quantitative
  Data**

  - [**Slides**](https://rpubs.com/PeneLope_PE/1217502){target="_blank"}

  - [**Notes**](https://rpubs.com/PeneLope_PE/1217503){target="_blank"}

  - [**PDF**](https://drive.google.com/file/d/1ugev0bGNZU1_fR4qUDkydQhEgG3hRTY5/view?usp=sharing){target="_blank"}

<br>

- **Lecture 4 - Data Visualizations, Boxplots, Outliers**

  - [**Slides**](https://rpubs.com/PeneLope_PE/1215503){target="_blank"}

  - [**Notes**](https://rpubs.com/PeneLope_PE/1215504){target="_blank"}

  - [**PDF**](https://drive.google.com/file/d/1o63tDFhCWLqsbHvsSp2o-yZ2FXAz58Py/view?usp=sharing){target="_blank"}

<br>

- **Lecture 3 - Measures of Variability**

  - [**Slides**](https://rpubs.com/PeneLope_PE/1215491){target="_blank"}

  - [**Notes**](https://rpubs.com/PeneLope_PE/1215492){target="_blank"}

  - [**PDF**](https://drive.google.com/file/d/1o5O5NY2gcX0oDGzniBb_cQJdGEZCRZPn/view?usp=sharing){target="_blank"}

<br>

- **Lecture 2 - Measures of Central Tendancy**

  - [**Slides**](https://rpubs.com/PeneLope_PE/1212300){target="_blank"}

  - [**Notes**](https://rpubs.com/PeneLope_PE/1212302){target="_blank"}

  - [**PDF**](https://drive.google.com/file/d/1B7X9kvma8Tfdrj1jkJmurekhhTH8rKll/view?usp=sharing){target="_blank"}

<br>

- **Lecture 1 - Course Introductions and Types of Data**

  - [**Slides**](https://rpubs.com/PeneLope_PE/1211541){target="_blank"}

  - [**Notes**](https://rpubs.com/PeneLope_PE/1211543){target="_blank"}

  - [**PDF**](https://drive.google.com/file/d/1mbjjq4OMmBb81ELnYVK73EVp0nIowyX8/view?usp=sharing){target="_blank"}

<br>

# Lectures 17-27

## Row {height="35%"}

<br>

**Linear relationships**

<br>

- **Lecture 27 - Introduction to Linear Transformations**

  - [**Slides**](https://rpubs.com/PeneLope_PE/1253624){target="_blank"}

  - [**Notes**](https://rpubs.com/PeneLope_PE/1253626){target="_blank"}

  - [**PDF**](https://drive.google.com/file/d/1Y48deDOVxyKr7PQWAHy2xgqM4ZNfn7-0/view?usp=sharing){target="_blank"}

<br>

- **Lecture 26 - Multiple Linear Regression**

  - [**Slides**](https://rpubs.com/PeneLope_PE/1252678){target="_blank"}

  - [**Notes**](https://rpubs.com/PeneLope_PE/1252679){target="_blank"}

  - [**PDF**](https://drive.google.com/file/d/1vVLQ4qVJBsQQgV0axE3lJAi8zbi9swQy/view?usp=sharing){target="_blank"}

<br>

- **Lecture 25 - Simple Linear Regression Continued**

  - [**Slides**](https://rpubs.com/PeneLope_PE/1247035){target="_blank"}

  - [**Notes**](https://rpubs.com/PeneLope_PE/1247033){target="_blank"}

  - [**PDF**](https://drive.google.com/file/d/1v-JYJh1Xx147UQUtK4JUJMnZoJnJwyhu/view?usp=sharing){target="_blank"}

<br>

- **Lecture 24 - Introduction to Simple Linear Regression**

  - [**Slides**](https://rpubs.com/PeneLope_PE/1245145){target="_blank"}

  - [**Notes**](https://rpubs.com/PeneLope_PE/1245147){target="_blank"}

  - [**PDF**](https://drive.google.com/file/d/1ztx45iKyfybZGIp8iI8NrWn_k5uHzk6N/view?usp=sharing){target="_blank"}

<br>

- **Lecture 21 - Introduction to Portfolio Management**

  - [**Slides**](https://rpubs.com/PeneLope_PE/1241207){target="_blank"}

  - [**Notes**](https://rpubs.com/PeneLope_PE/1241208){target="_blank"}

  - [**PDF**](https://drive.google.com/file/d/1-IfsrZvPQ9vAZsGdJWSzhQaMbGVdfBwC/view?usp=sharing){target="_blank"}

<br>

**Hypotheses and Correlations**

- **Lecture 20 - Introduction to Correlation and
  Covariance**

  - [**Slides**](https://rpubs.com/PeneLope_PE/1239229){target="_blank"}

  - [**Notes**](https://rpubs.com/PeneLope_PE/1239230){target="_blank"}

  - [**PDF**](https://drive.google.com/file/d/1CvWpRPqul2sx2PI121_ryCa0P_cV8jRc/view?usp=sharing){target="_blank"}

<br>

- **Lecture 19 - Contingency Tables and Two Sample Tests of
  Proportions**

  - [**Slides**](https://rpubs.com/PeneLope_PE/1238251){target="_blank"}

  - [**Notes**](https://rpubs.com/PeneLope_PE/1238253){target="_blank"}

  - [**PDF**](https://drive.google.com/file/d/1Ra-ivkv6cN8UlZLvQCpKjzEDGzn0c9tD/view?usp=sharing){target="_blank"}

  - [**Posit Cloud**](https://posit.cloud/){target="_blank"}

<br>

- **Lecture 18 - Two Sample t-Tests and Confidence
  Intervals**

  - [**Slides**](https://rpubs.com/PeneLope_PE/1236174){target="_blank"}

  - [**Notes**](https://rpubs.com/PeneLope_PE/1236179){target="_blank"}

  - [**PDF**](https://drive.google.com/file/d/1bxnZjFgGjISe0zjIeja2Kmr4Lx7EKWa4/view?usp=sharing){target="_blank"}

<br>

- **Lecture 17 - Language of Hypothesis Testing/One Sample
  t-tests**

  - [**Slides**](https://rpubs.com/PeneLope_PE/1235312){target="_blank"}

  - [**Notes**](https://rpubs.com/PeneLope_PE/1235311){target="_blank"}

  - [**PDF**](https://drive.google.com/file/d/1XujkJNFOnC7viS_VqTQA_6Bl09RradHz/view?usp=sharing){target="_blank"}

# Installing R and Rstudio

## Row

**Notes**

- Students in MAS 261 are NOT required to install R and
  RStudio on their laptop but they are welcome to.

- R and RStudio are two separate software components.
  Download and install R and then download and install
  RStudio.

- When updating your version of R or RStudio, uninstall the
  previous version first.

## Row

**Downloading and Installing R**

- R can be downloaded from the [**Comprehensive R Archive
  Network**](https://cran.r-project.org/){target="_blank"}

- [**Video Demo for downloading and installing R on a
  Windows
  OS**](https://video.syr.edu/media/t/1_m775pb4g/352723432){target="_blank"}

- Mac installation is similar.

- If you do not need to uninstall an out-of-date version of
  R, skip to 0:53.

- The current version of R is **`4.6.1`**, which was
  released on June 24, 2026.

<br>

**Downloading and Installing RStudio**

- RStudio can be downloaded from
  [**Posit**](https://posit.co/download/rstudio-desktop/){target="_blank"},
  the parent company of RStudio.

- [**Video Demo for downloading and installing RStudio on a
  Windows
  OS**](https://video.syr.edu/media/t/1_my39umi3/352723432){target="_blank"}

- Mac installation is similar.

- If you do not need to uninstall an out-of-date version of
  RStudio, skip to 0:56.

- The version shown in the video is from August of 2024 and
  is out-of-date.

- The current version of RStudio is **`2026.8.1.195`**,
  which was published on August 18, 2026.
